---
title: "Movie Dataset — Exploratory Data Analysis"
author: "Group 01"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 5
)
```

---

# Introduction

This report presents an Exploratory Data Analysis (EDA) on an IMDB-style movies dataset. The goal is to uncover patterns and relationships between movie features such as budget, gross revenue, IMDb scores, genres, countries, and release years.

---

# Setup & Libraries

```{r libraries, eval=FALSE}
install.packages(c("readr", "dplyr", "ggplot2", "gridExtra",
                   "GGally", "corrplot", "plotly"))
```

```{r load-libraries}
library(readr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(corrplot)
library(plotly)
```

---

# Load Data

The dataset is loaded directly from Google Drive using its file ID.

```{r load-data}
FILE_ID <- "1mASYC6eyEajnigyR6F52fqd9bhG18BRP"
url <- paste0("https://drive.google.com/uc?export=download&id=", FILE_ID)
movies <- read_csv(url)
```

---

# Data Cleaning

## Inspect Missing Values

```{r missing-check}
movies[!complete.cases(movies), ]
colSums(is.na(movies))
```

## Remove Irrelevant Columns

We drop Facebook-related engagement columns, actor names, aspect ratio, and movie links as they are not useful for our analysis.

```{r drop-columns}
movies <- movies %>%
  select(-director_facebook_likes, -actor_3_facebook_likes,
         -actor_1_facebook_likes, -actor_3_name, -actor_2_facebook_likes,
         -cast_total_facebook_likes, -aspect_ratio,
         -movie_facebook_likes, -movie_imdb_link, -actor_2_name,
         -facenumber_in_poster, -actor_3_name)
```

## Remove Extreme Budget Outlier

```{r remove-max-budget}
movies <- movies %>%
  filter(budget != max(budget, na.rm = TRUE))
```

## Feature Engineering

We create million-scale financial variables and extract the primary genre from the pipe-separated genre string.

```{r feature-engineering}
movies <- movies %>%
  mutate(
    budget_million  = budget / 1e6,
    gross_million   = gross / 1e6,
    profit_million  = (gross - budget) / 1e6,
    main_genre      = sapply(strsplit(genres, "\\|"), `[`, 1)
  )
```

## Impute Missing Values

**Color** — imputed with mode:

```{r impute-color}
mode_color <- names(which.max(table(movies$color)))
movies$color[is.na(movies$color)] <- mode_color
sum(is.na(movies$color))
```

**Duration** — imputed with mean:

```{r impute-duration}
mean_duration <- mean(movies$duration, na.rm = TRUE)
movies$duration[is.na(movies$duration)] <- round(mean_duration)
sum(is.na(movies$duration))
```

**User & Critic Reviews** — imputed with mean:

```{r impute-reviews}
movies$num_user_for_reviews[is.na(movies$num_user_for_reviews)] <-
  mean(movies$num_user_for_reviews, na.rm = TRUE)

movies$num_critic_for_reviews[is.na(movies$num_critic_for_reviews)] <-
  mean(movies$num_critic_for_reviews, na.rm = TRUE)
```

## Remove Rows Missing Budget or Gross

```{r remove-na-financial}
movies <- movies %>%
  filter(!is.na(gross), !is.na(budget))

sum(is.na(movies$gross))
sum(is.na(movies$budget))
```

## Dataset Summary

```{r summary}
summary(movies)
```

---

# Univariate Analysis

## IMDb Score Distribution

```{r imdb-hist}
ggplot(movies, aes(x = imdb_score)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Distribution of IMDb Scores", x = "IMDb Score", y = "Count")
```

## Movie Duration Distribution

```{r duration-hist}
ggplot(movies, aes(x = duration)) +
  geom_histogram(binwidth = 10, fill = "lightgreen", color = "black") +
  labs(title = "Distribution of Movie Duration", x = "Duration (minutes)", y = "Count")
```

## Budget, Gross & Profit Distributions

```{r financial-hist}
ggplot(movies, aes(x = budget_million)) +
  geom_histogram(binwidth = 500, fill = "red", color = "black") +
  labs(title = "Distribution of Movie Budgets", x = "Budget (USD Million)", y = "Count")

ggplot(movies, aes(x = gross_million)) +
  geom_histogram(binwidth = 500, fill = "red", color = "black") +
  labs(title = "Distribution of Movie Gross Revenue", x = "Gross (USD Million)", y = "Count")

ggplot(movies, aes(x = profit_million)) +
  geom_histogram(binwidth = 500, fill = "red", color = "black") +
  labs(title = "Distribution of Movie Profits", x = "Profit (USD Million)", y = "Count")
```

## Color Format Distribution

```{r color-pie}
movies %>%
  count(color) %>%
  mutate(prop  = n / sum(n),
         label = paste0(color, " (", round(prop * 100, 1), "%)")) %>%
  ggplot(aes(x = "", y = prop, fill = color)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5)) +
  labs(title = "Distribution of Movie Colors") +
  theme_void() +
  scale_fill_brewer(palette = "Set2")
```

## Movies by Genre

```{r genre-bar}
genre_counts <- movies %>% count(main_genre)

ggplot(genre_counts, aes(x = reorder(main_genre, n), y = n)) +
  geom_col(fill = "orange") +
  coord_flip() +
  labs(title = "Number of Movies by Genre", x = "Genre", y = "Number of Movies")
```

## Top 10 Languages

```{r language-bar}
movies %>%
  count(language) %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = reorder(language, n), y = n)) +
  geom_col(fill = "cyan") +
  coord_flip() +
  labs(title = "Top 10 Languages by Number of Movies", x = "Language", y = "Number of Movies")
```

## Top 10 Countries by Production

```{r country-bar}
movies %>%
  count(country) %>%
  top_n(10, n) %>%
  ggplot(aes(x = reorder(country, n), y = n)) +
  geom_col(fill = "gold") +
  coord_flip() +
  labs(title = "Top 10 Countries by Movie Production", x = "Country", y = "Number of Movies")
```

## Density Plots

```{r density-plots}
ggplot(movies, aes(x = imdb_score)) +
  geom_density(fill = "skyblue", alpha = 0.6) +
  labs(title = "Density Plot of IMDb Scores", x = "IMDb Score", y = "Density")

ggplot(movies, aes(x = duration)) +
  geom_density(fill = "lightgreen", alpha = 0.6) +
  labs(title = "Density Plot of Movie Duration", x = "Duration (minutes)", y = "Density")

movies %>%
  filter(title_year >= 1980) %>%
  ggplot(aes(x = title_year)) +
  geom_density(fill = "purple", alpha = 0.6) +
  labs(title = "Density Plot of Movie Release Years (1980–Present)",
       x = "Release Year", y = "Density")
```

---

# Bivariate Analysis

## Budget vs Gross Revenue

```{r budget-vs-gross}
ggplot(movies, aes(x = budget_million, y = gross_million)) +
  geom_point(alpha = 0.5, color = "blue") +
  labs(title = "Budget vs Gross Revenue",
       x = "Budget (USD Million)", y = "Gross Revenue (USD Million)")
```

## Duration vs IMDb Score

```{r duration-vs-imdb}
ggplot(movies, aes(x = duration, y = imdb_score)) +
  geom_point(alpha = 0.5, color = "darkgreen") +
  labs(title = "Duration vs IMDb Score", x = "Duration (minutes)", y = "IMDb Score")
```

## Financial Metrics vs IMDb Score

```{r financial-vs-imdb}
p1 <- ggplot(movies, aes(x = budget_million, y = imdb_score)) +
  geom_point(alpha = 0.5, color = "blue") +
  labs(title = "IMDb Score vs Budget", x = "Budget (Million USD)", y = "IMDb Score") +
  theme_minimal()

p2 <- ggplot(movies, aes(x = gross_million, y = imdb_score)) +
  geom_point(alpha = 0.5, color = "green") +
  labs(title = "IMDb Score vs Gross", x = "Gross Revenue (Million USD)", y = "IMDb Score") +
  theme_minimal()

p3 <- ggplot(movies, aes(x = profit_million, y = imdb_score)) +
  geom_point(alpha = 0.5, color = "red") +
  labs(title = "IMDb Score vs Profit", x = "Profit (Million USD)", y = "IMDb Score") +
  theme_minimal()

grid.arrange(p1, p2, p3, ncol = 3)
```

## IMDb Score & Duration by Genre (Boxplots)

```{r genre-boxplots}
ggplot(movies, aes(x = main_genre, y = imdb_score)) +
  geom_boxplot(fill = "pink") +
  coord_flip() +
  labs(title = "IMDb Scores by Genre", x = "Genre", y = "IMDb Score")

ggplot(movies, aes(x = main_genre, y = duration)) +
  geom_boxplot(fill = "violet") +
  coord_flip() +
  labs(title = "Movie Duration by Genre", x = "Genre", y = "Duration (minutes)")
```

## Top 5 Genres — Pie Chart

```{r genre-pie}
movies %>%
  mutate(main_genre = sapply(strsplit(genres, "\\|"), `[`, 1)) %>%
  count(main_genre) %>%
  top_n(5, n) %>%
  ggplot(aes(x = "", y = n, fill = main_genre)) +
  geom_col() +
  coord_polar(theta = "y") +
  labs(title = "Proportion of Top 5 Genres") +
  theme_void()
```

---

# Country & Language Analysis

```{r country-setup}
top_countries  <- movies %>% count(country)  %>% slice_max(n, n = 10) %>% pull(country)
top_languages  <- movies %>% count(language) %>% slice_max(n, n = 10) %>% pull(language)
top_5_countries <- movies %>% count(country)  %>% slice_max(n, n = 5)  %>% pull(country)
top_5_languages <- movies %>% count(language) %>% slice_max(n, n = 5)  %>% pull(language)
```

## IMDb Score, Budget & Profit by Top Countries

```{r country-boxplots}
p1 <- movies %>%
  filter(country %in% top_countries) %>%
  ggplot(aes(x = country, y = imdb_score)) +
  geom_boxplot(fill = "lightgreen") + coord_flip() +
  labs(title = "IMDb Score by Country", x = "Country", y = "IMDb Score") + theme_minimal()

p2 <- movies %>%
  filter(country %in% top_countries) %>%
  ggplot(aes(x = country, y = budget_million)) +
  geom_boxplot(fill = "skyblue") + coord_flip() +
  labs(title = "Budget by Country", x = "Country", y = "Budget (Million USD)") + theme_minimal()

p3 <- movies %>%
  filter(country %in% top_countries) %>%
  ggplot(aes(x = country, y = profit_million)) +
  geom_boxplot(fill = "orange") + coord_flip() +
  labs(title = "Profit by Country", x = "Country", y = "Profit (Million USD)") + theme_minimal()

grid.arrange(p1, p2, p3, ncol = 1)
```

## IMDb Score by Top Languages

```{r language-boxplot}
movies %>%
  filter(language %in% top_languages) %>%
  ggplot(aes(x = language, y = imdb_score)) +
  geom_boxplot(fill = "lightblue") +
  coord_flip() +
  labs(title = "IMDb Score by Top 10 Languages", x = "Language", y = "IMDb Score") +
  theme_minimal()
```

## Genre Distribution by Country & Language

```{r genre-country-language}
movies %>%
  filter(country %in% top_5_countries) %>%
  count(main_genre, country) %>%
  ggplot(aes(x = main_genre, y = n, fill = country)) +
  geom_col(position = "dodge") +
  labs(title = "Movies by Genre and Country", x = "Genre", y = "Number of Movies") +
  theme_minimal()

movies %>%
  filter(language %in% top_5_languages) %>%
  count(main_genre, language) %>%
  ggplot(aes(x = main_genre, y = n, fill = language)) +
  geom_col(position = "dodge") +
  labs(title = "Movies by Genre and Language", x = "Genre", y = "Number of Movies") +
  theme_minimal()

movies %>%
  count(main_genre, color) %>%
  ggplot(aes(x = main_genre, y = n, fill = color)) +
  geom_col(position = "dodge") +
  labs(title = "Movie Color Format by Genre", x = "Genre", y = "Number of Movies") +
  theme_minimal()
```

---

# Multivariate Analysis

## Budget vs Gross/Profit Colored by Genre

```{r multivariate-genre}
ggplot(movies, aes(x = budget_million, y = gross_million, color = main_genre)) +
  geom_point(alpha = 0.6) +
  labs(title = "Budget vs Gross (Colored by Genre)",
       x = "Budget (Million USD)", y = "Gross Revenue (Million USD)") +
  theme_minimal()

ggplot(movies, aes(x = duration, y = imdb_score, color = country)) +
  geom_point(alpha = 0.6) +
  labs(title = "Duration vs IMDb Score (Colored by Country)",
       x = "Duration (Minutes)", y = "IMDb Score") +
  theme_minimal()

ggplot(movies, aes(x = profit_million, y = imdb_score, color = language)) +
  geom_point(alpha = 0.6) +
  labs(title = "Profit vs IMDb Score (Colored by Language)",
       x = "Profit (Million USD)", y = "IMDb Score") +
  theme_minimal()
```

## Regression Lines: Budget & Duration vs Profit by Genre

```{r regression-genre}
movies$main_genre <- as.factor(movies$main_genre)

p1 <- ggplot(movies, aes(x = budget_million, y = gross_million, color = main_genre)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Budget vs Gross by Genre") +
  xlab("Budget (Million USD)") + ylab("Gross (Million USD)") + theme_minimal()

p2 <- ggplot(movies, aes(x = duration, y = profit_million, color = main_genre)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Duration vs Profit by Genre") +
  xlab("Duration (Minutes)") + ylab("Profit (Million USD)") + theme_minimal()

p3 <- ggplot(movies, aes(x = imdb_score, y = profit_million, color = main_genre)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE) +
  ggtitle("IMDb Score vs Profit by Genre") +
  xlab("IMDb Score") + ylab("Profit (Million USD)") + theme_minimal()

grid.arrange(p1, p2, p3, ncol = 1)
```

---

# Temporal Trends

```{r yearly-trends}
yearly_stats <- movies %>%
  group_by(title_year) %>%
  summarise(
    avg_imdb_score     = mean(imdb_score, na.rm = TRUE),
    avg_budget_million = mean(budget_million, na.rm = TRUE),
    avg_profit_million = mean(profit_million, na.rm = TRUE),
    movies_count       = n()
  ) %>%
  arrange(title_year)

p1 <- ggplot(yearly_stats, aes(x = title_year, y = avg_imdb_score)) +
  geom_line(color = "blue") + geom_point() +
  labs(title = "Average IMDb Score per Year", x = "Year", y = "Avg IMDb Score")

p2 <- ggplot(yearly_stats, aes(x = title_year, y = avg_budget_million)) +
  geom_line(color = "green") + geom_point() +
  labs(title = "Average Budget per Year", x = "Year", y = "Avg Budget (Million)")

p3 <- ggplot(yearly_stats, aes(x = title_year, y = avg_profit_million)) +
  geom_line(color = "purple") + geom_point() +
  labs(title = "Average Profit per Year", x = "Year", y = "Avg Profit (Million)")

p4 <- ggplot(yearly_stats, aes(x = title_year, y = movies_count)) +
  geom_line(color = "red") + geom_point() +
  labs(title = "Number of Movies Released per Year", x = "Year", y = "Count")

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

---

# Top Movies & Actors

## Top 10 IMDb Rated Movies

```{r top-movies}
movies %>%
  arrange(desc(imdb_score)) %>%
  slice(1:10) %>%
  ggplot(aes(x = reorder(movie_title, imdb_score), y = imdb_score)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  labs(title = "Top 10 IMDb Rated Movies", x = "Movie Title", y = "IMDb Score")
```

## Top 10 Actors by Average Profit

```{r top-actors}
top_actors_profit <- movies %>%
  group_by(actor_1_name) %>%
  summarise(avg_profit = mean(profit_million, na.rm = TRUE)) %>%
  arrange(desc(avg_profit)) %>%
  slice_head(n = 10)

ggplot(top_actors_profit, aes(x = reorder(actor_1_name, avg_profit), y = avg_profit)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Actors by Average Profit", x = "Actor", y = "Average Profit (Million USD)")
```

---

# Correlation Analysis

## Pair Plot

```{r pair-plot, fig.height=7}
movies$color_num <- as.numeric(as.factor(movies$main_genre))

pairs(
  movies[, c("budget_million", "gross_million", "profit_million", "duration")],
  col  = movies$color_num,
  pch  = 19,
  main = "Pair Plot of Selected Movie Features by Genre"
)
```

## Correlation Matrix

```{r corrplot, fig.height=8}
numeric_movies <- movies %>% select(where(is.numeric))
cor_matrix <- cor(numeric_movies, use = "complete.obs")

corrplot(cor_matrix, method = "color", type = "upper",
         tl.col = "black", tl.srt = 45,
         addCoef.col = "black",
         number.cex = 0.7)
```

## Interactive Correlation Heatmap

```{r plotly-heatmap}
plot_ly(
  x      = colnames(cor_matrix),
  y      = colnames(cor_matrix),
  z      = cor_matrix,
  type   = "heatmap",
  colors = colorRamp(c("red", "white", "blue")),
  zmin   = -1, zmax = 1
) %>%
  layout(
    title  = "Correlation Heatmap of Movies Dataset",
    xaxis  = list(title = ""),
    yaxis  = list(title = "", autorange = "reversed")
  )
```

---

# Outlier Analysis

## Define Outlier Removal Function

```{r outlier-function}
remove_outliers <- function(x) {
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR_val <- Q3 - Q1
  lower <- Q1 - 1.5 * IQR_val
  upper <- Q3 + 1.5 * IQR_val
  x[x < lower | x > upper] <- NA
  return(x)
}

movies$budget_clean     <- remove_outliers(movies$budget_million)
movies$gross_clean      <- remove_outliers(movies$gross_million)
movies$profit_clean     <- remove_outliers(movies$profit_million)
movies$imdb_score_clean <- remove_outliers(movies$imdb_score)
```

## Before vs After Outlier Removal

```{r outlier-plots, fig.height=12}
p_bg_before <- ggplot(movies, aes(x = budget_million, y = gross_million)) +
  geom_point(alpha = 0.4, color = "grey30") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Budget vs. Gross (Before)", x = "Budget", y = "Gross") + theme_minimal()

p_bg_after <- ggplot(movies, aes(x = budget_clean, y = gross_clean)) +
  geom_point(alpha = 0.4, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Budget vs. Gross (After)", x = "Budget (Cleaned)", y = "Gross (Cleaned)") + theme_minimal()

p_bp_before <- ggplot(movies, aes(x = budget_million, y = profit_million)) +
  geom_point(alpha = 0.4, color = "grey30") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Budget vs. Profit (Before)", x = "Budget", y = "Profit") + theme_minimal()

p_bp_after <- ggplot(movies, aes(x = budget_clean, y = profit_clean)) +
  geom_point(alpha = 0.4, color = "green") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Budget vs. Profit (After)", x = "Budget (Cleaned)", y = "Profit (Cleaned)") + theme_minimal()

p_ip_before <- ggplot(movies, aes(x = imdb_score, y = profit_million)) +
  geom_point(alpha = 0.4, color = "grey30") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "IMDb Score vs. Profit (Before)", x = "IMDb Score", y = "Profit") + theme_minimal()

p_ip_after <- ggplot(movies, aes(x = imdb_score_clean, y = profit_clean)) +
  geom_point(alpha = 0.4, color = "purple") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "IMDb Score vs. Profit (After)", x = "IMDb Score (Cleaned)", y = "Profit (Cleaned)") + theme_minimal()

grid.arrange(p_bg_before, p_bg_after,
             p_bp_before, p_bp_after,
             p_ip_before, p_ip_after,
             ncol = 2)
```

---

# Conclusion

This EDA revealed several key findings from the movies dataset:

- **IMDb scores** are roughly normally distributed, centered around 6–7.
- **Budget and gross revenue** show a positive correlation — higher budgets tend to yield higher revenue.
- **Genre** plays a significant role in both IMDb scores and financial performance.
- **The USA** dominates movie production, with English being the most common language.
- **Drama and Comedy** are the most frequently produced genres.
- **Outlier removal** improved the clarity of regression relationships significantly.
- **Yearly trends** indicate rising budgets over the decades with fluctuating average scores.
